The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
Methods updates

% Set some cref options
\newcommand{\crefrangeconjunction}{--} % Change crefrange from "to" to dashed
\crefname{figure}{Figure}{Figure}

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
In addition to photoconversion experiments, we carried out particle tracking inside the brain ventricles of a larval zebrafish, to demonstrate the motile-cilia and cardiac-cycle related flow features observed and described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}. A schematical illustration of the flow features in the brain ventricles is shown in~\cref{fig:figure1}, panel \textbf{a}. Our experimental setup for the particle tracking is identical to the ones considered there (\lyng{HH: Is this true?}), and we refer to that paper for further details than the ones provided here in the following.

@@

\subsection*{Image-based computational geometries of zebrafish brain ventricles}
We generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. From the imaging data, we constructed a computational mesh of the ventricles (\cref{fig:figure1}, panel \textbf{d}). The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic  (anterior), the di-/mesencephalic  (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. Parts of the ventricular surfaces were marked to be lined with motile cilia~\cite{Olstad2019CiliaryDevelopment} (\cref{fig:figure1}d, pink markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

%\subsection*{Computational geometry variations}
The morphology and geometry of the brain ventricles vary across zebrafish individuals, and under physiological and pathological conditions. Motivated by an interest in how geometry impacts solute distribution, we also consider four variations in the ventricular geometry: by shrinking the fore-mid brain connection, the middle ventricle, the mid-hind brain connection, and all three simultaneously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry. \mer{MER: More specifics (i.e.~1 sentence and/or inserted info) here on how much these were shrunk and how.}

%% The brain ventricular geometry varies significantly across zebrafish individuals, both healthy and sick animals. Motivated by an interest in how these variations can impact solute transport in the brain ventricles, we study the impact on simulation results when modifying the computational geometry and mesh in four different scenarios. Because our computational mesh is inflated by the dye-injection performed before image segmentation, we study four different mesh versions where we shrink: the fore-mid brain connection; the middle ventricle; the mid-hind brain connection; and all of the three above simultaenously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.

\subsection*{Computational CSF dynamics}

The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment} within the brain ventricles (\cref{fig:figure1}, panels \textbf{a}--\textbf{c}). In addition, cardiac pulsations induces pulsatile CSF flow. We model this flow of CSF in the brain ventricles by the time-dependent, incompressible Stokes equations which read as follows. Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$ and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,
      \label{eq:stokes_eq_mom}\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions. Furthermore, $\rho$ is the CSF density and $\mu$ is the dynamic viscosity of the CSF. We introduce the tangential traction $\bsigpar$ as the tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$, and $P_{\nn}(\rr)$ is the tangential  projection of a vector $\rr$ onto $\Gamma$. We consider three types of boundary conditions:  (i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$), (ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$), and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$). Unless noted otherwise, the free-slip condition is applied on the boundary. We assume that the system starts at rest ($\uu(t = 0) = \mathbf{0}$).

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\cref{fig:figure1}, panel \textbf{d}), we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces of the cilia acting on the CSF. We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$ and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the cilia region (see~\cref{subsec:appendixA1}). Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by model calibration against the experimental data~\cite{Olstad2019CiliaryDevelopment}.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and $\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts of the boundary $\Gp$~(\cref{fig:figure1}, panel \textbf{d}, blue markers). The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow, and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. The angular cardiac frequency $\omega=2\pi f=6.97$ rad/s was set based on a measured cardiac frequency of $f=2.22$ Hz.

%% \iffalse
%% \subsection*{Computational cerebrospinal fluid dynamics in a moving reference frame}
%%  An Arbitrary Lagrangian Eulerian (ALE) formulation~\cite{san2009convergence} is used as an alternative to the normal pressure boundary conditions to model back-and-forth flow induced by cardiac pulsations. In a reference frame moving with a velocity $\uuref=(A_{\mathrm{ref}}\sin{\omega t}, 0, 0)$, we reformulate the Stokes equations~\eqref{eq:stokes_eqs} using an ALE formulation:
%% \begin{subequations}
%%     \begin{alignat}{2}
%%         \rho\bigg(\pdifft{\uu} + (\uu_{\mathrm{ref}}\cdot\nabla)\uu\bigg) &= \nabla\cdot\bm{\sigma}(\uu, p), \label{eq:mom_eq_ALE}\\
%%         \nabla\cdot\uu &= 0.
%%     \end{alignat}
%%     \label{eq:stokes_eqs_ALE}%
%% \end{subequations}%
%% We then solve for the velocity $\mathbf{w} = \uu - \uuref$ in the moving reference frame, where $\uu$ is the velocity in the fixed reference frame. The amplitude $A_{\mathrm{ref}}=919$ $\mu$m/s, calibrated based on experimental data, determines the magnitude of the pulsatile flow, and $\omega=2\pi f$ is the angular cardiac frequency. In the ALE formulation, we consider the impermeability boundary condition $\mathbf{w}\cdot\nn=0$ in the normal direction and the free-slip boundary condition $\bsigpar(\mathbf{w}, p) = \mathbf{0}$ in the tangential direction.
%% \fi

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$. The ventricular walls are assumed to be impermeable (admitting \emph{no flux} i.e.~$\JJ\cdot\nn=0$), except at the anterior/posterior boundary $\Gamma_p$. \fixme{At the anterior/posterior boundary we allow the solute concentration to be advected into and out of the domain in a manner that ensures conservation of mass (see \Cref{sec:appendix:X} for details).}
\mer{MER: @Halvor: could you have a go at moving the precise description to the Appendix?}
\fixme{There, we consider either prescribed \emph{influx} in the form of a time-periodic condition on the total flux $(\JJ\cdot\nn)|_{\mathrm{in}}^{n+1}=(\JJ\cdot\nn)|_{\mathrm{out}}^{n}$ and/or \emph{outflux} with zero diffusive flux $D\nabla c\cdot\nn=0$.}
\fixme{Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$; $\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep. This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero.}
%\mer{MER: I moved this paragraph from the section below, because it seemed to belong here. But needs edits.}

%Unless stated otherwise, the no-flux condition $\JJ\cdot\nn=0$ was applied on the boundary.
%(ii) \emph{injection site} (imposing the Dirichlet boundary condition $c=\Tilde{c}$),

%\subsection*{Modeling protein photoconversion}

@@

%To represent the solute injection at $\Ginj$ (cf.~\cref{fig:injection_sites}), we impose a constant Dirichlet boundary condition $c=\Tilde{c}$. Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$; $\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep. This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero. 
%\lyng{MOVE? Following previous works~\cite{papanastasiou1992new, griffiths1997no, lynch2020numerical}, the advective flux $c\uu$ on $\Gout$ was left as an unknown in the variational form.}

\subsection*{Estimation of diffusion coefficients via the Stokes-Einstein relation}
%\label{subsec:diff_coeffs}

Motivated by an interest in molecules that are significant in neural development, we study the transport of molecules resembling Dendra2 Fluorescent Protein (Dendra2), Starmaker+Green Fluorescent Protein (STM+GFP) and Exosomes. To estimate the diffusion coefficients of these, when and if not available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R} ,
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}. In~\eqref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant, $T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the solvent (in this case the CSF), and $R$ is the radius of the molecule. For the exosomes, we calculate $D$ using~\eqref{eq:D_stokes_einstein} with $R=200$ nm. This radius covers the slowest diffusing exosomes, since the smaller exosomes would have diffusion coefficients similar to the other molecules. For STM+GFP, we estimated the diffusion coefficient by extrapolating the experimentally observed diffusion coefficient of GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that if $D$ scales with $1/R$ according to~\eqref{eq:D_stokes_einstein}, $D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}. The resulting diffusion coefficients used in the simulations are summarized in \cref{tab:diff_coeff_table}.

%% In summary, in the simulations, we use the diffusion coefficients $D_1$, $D_2$ and $D_3$ (\cref{tab:parameter_table}) for these three molecules

@@

%% The molecular diffusion coefficients were estimated in part based on values reported in the literature, and in part via extrapolation with the Stokes--Einstein relation (\cref{tab:diff_coeff_table}). The Stokes-Einstein relation
\begin{table}[h!]
    \centering
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM+GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes. STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively. (*) The value was extrapolated based on the value of $D$ for GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that $D$ scales with the cube root of the molecule mass. (\textdagger) The value was estimated using the Stokes-Einstein relation~\eqref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
\end{table}

%% The diffusion coefficient was estimated through extrapolation of previous measurements~\cite{GuraSadovsky2017MeasurementExpansion, Mahmood2023ExosomeTemperature, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} via the Stokes-Einstein relation

\subsection*{Physical parameters}
%\label{subsec:parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion~(\cref{sec:appC_particle_transport}) was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you introduce parameter values when introduced, and move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Then remove this section and \Cref{tab:parameter_table}.}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

>>>>

% Set some cref options
\newcommand{\crefrangeconjunction}{--} % Change crefrange from "to" to dashed

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
In addition to photoconversion experiments, we carried out particle tracking inside the brain ventricles of a larval zebrafish, to demonstrate the motile-cilia and cardiac-cycle related flow features observed and described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}. A schematical illustration of the flow features in the brain ventricles is shown in~\Cref{fig:figure1}a. Our experimental setup for the particle tracking is identical to the ones considered there (\lyng{HH: Is this true?}), and we refer to that paper for further details than the ones provided here in the following.

@@

\subsection*{Image-based computational geometries of zebrafish brain ventricles}
We generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:figure1}d). The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic  (anterior), the di-/mesencephalic  (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:figure1}d, pink markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

The morphology and geometry of the brain ventricles vary across zebrafish individuals, and under physiological and pathological conditions. 
Motivated by an interest in how geometry impacts solute distribution, we also consider four variations in the ventricular geometry: by shrinking the fore-mid brain connection, 
the middle ventricle, the mid-hind brain connection, and all three simultaneously. 
We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry. \mer{MER: More specifics (i.e.~1 sentence and/or inserted info) here on how much these were shrunk and how.}

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment} within the brain ventricles (\Cref{fig:figure1}a--c).
In addition, cardiac pulsations induces pulsatile CSF flow. We model this flow of CSF in the brain ventricles by the time-dependent, incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$ and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,
      \label{eq:stokes_eq_mom}\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions. Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density and $\mu=6.97\times 10^{-4}$ Pa is the dynamic viscosity of the CSF.
We introduce the tangential traction $\bsigpar$ as the tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$, and $P_{\nn}(\rr)$ is the tangential  projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:  (i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the boundary. We assume that the system starts at rest ($\uu(\xx, t=0) = \mathbf{0}$).

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:figure1}d), we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$ and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by model calibration against the experimental data~\cite{Olstad2019CiliaryDevelopment}.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and $\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts of the boundary $\Gp$~(\Cref{fig:figure1}d, blue markers).
The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow, and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
The angular cardiac frequency $\omega=2\pi f=6.97$ rad/s was set based on a measured cardiac frequency of $f=2.22$ Hz.

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux} i.e.~$\JJ\cdot\nn=0$), except at the anterior/posterior boundary $\Gamma_p$.
\fixme{At the anterior/posterior boundary we allow the solute concentration to be advected into and out of the domain in a manner that ensures conservation of mass (see~\Cref{sec:appendix:X} for details).}
\mer{MER: @Halvor: could you have a go at moving the precise description to the Appendix?}
\fixme{There, we consider either prescribed \emph{influx} in the form of a time-periodic condition on the total flux $(\JJ\cdot\nn)|_{\mathrm{in}}^{n+1}=(\JJ\cdot\nn)|_{\mathrm{out}}^{n}$ and/or \emph{outflux} with zero diffusive flux $D\nabla c\cdot\nn=0$.}
\fixme{Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$;
$\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep.
This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero.}

@@

\subsection*{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development, we study the transport of molecules resembling Dendra2 Fluorescent Protein (Dendra2), Starmaker+Green Fluorescent Protein (STM+GFP) and Exosomes.
To estimate the diffusion coefficients of these, when and if not available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\eqref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant, $T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the solvent (in this case the CSF), and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\eqref{eq:D_stokes_einstein} with $R=200$ nm. This radius covers the slowest diffusing exosomes, since the smaller exosomes would have diffusion coefficients similar to the other molecules.
For STM+GFP, we estimated the diffusion coefficient by extrapolating the experimentally observed diffusion coefficient of GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\eqref{eq:D_stokes_einstein}, $D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients used in the simulations are summarized in \Cref{tab:diff_coeff_table}.

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM+GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes. STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass. (\textdagger) The value was estimated using the Stokes-Einstein relation~\eqref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
\end{table}

\subsection*{Physical parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\Cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you introduce parameter values when introduced, and move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Then remove this section and \Cref{tab:parameter_table}.}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

\subsection*{Computational flow modelling scenarios}
To study the impact of the motile cilia and the cardiac pulsations on the CSF flow, we consider the following computational flow scenarios, each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

    \item A \emph{cilia-only/no-cardiac} flow model including contributions from the cilia forces but no cardiac pulsations.
    Here, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$, and \mer{XXX on the anterior/posterior boundary $\Gp$.} % Model I

    \item A \emph{cardiac-only/no-cilia} flow model including contributions from the cardiac pulsations but no cilia forces.
    Here, we set the tangential traction to zero on the cilia boundary $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$.  % Model II
\end{itemize}
